THESIS PROJECT PROPOSAL

### 2019-11-01 project_ashish/paper.tex
Update paper.tex

The standard model (SM) of particle physics is so far the best theoretical model to describe the interaction of elementary 
particles using three of the four fundamental forces of nature which are electromagnetic force, strong nuclear force and the weak
nuclear force. Gravitational force is neglected as the strength of this force is very weak at the scales over which elementary 
particle interact with each other. The standard model (SM) of particle physics is divided into two categories, bosonic sector 
and fermionic sector. Bosonic sector contain particles called bosons which mediate the fundamental forces of nature and the
fermionic sector contain particles called fermions which make up all the matter in our universe. SM has three generations of 
matter (fermions) particles.The first generation of fermions consists of up(u) quark, down(d) quark, electron and electron 
neutrino,second generation consist of charm(c) quark, strange(s) quark, muon and muon neutrino and the third generation of matter 
particles has top(t) quark,bottom(b) quark, tau and tau neutrino. The bosonic sector consist of gauge bosons like gluon, photon, 
W^+ W^- Z^0 which mediate strong force, electromagnetic force and weak force respectively. There is one more particle in the 
standard model called the Higgs Boson which gives mass to SM particles via electroweak symmetry breaking mechanism. Higgs boson 
can be produced at the particle colliders like the Large Hadron Collider(LHC) in Geneva,Switzerland.

There are various production modes of Higgs boson like the gluon gluon fusion(ggF), vector boson fusion(VBF), Higgs production 
in association with vector boson(VH, V=W or Z), Higgs production with top quark and top anti-quark(ttH) and Higgs production with 
a single top quark and a quark jet(tHq). Each production channel has its own importance to probe the properties and the coupling 
strength of Higgs boson to fermions and bosons. The main production mode of Higgs bosons at LHC is gluon gluon fusion(ggF). As 
bosons like gluons and photons are massless, they do not interact directly with the Higgs boson and processes like ggF or Higgs 
boson decay to pair of photons are not possible at tree level and can only proceed via loop diagrams which involve W boson or top 
quark in the loop. The Higgs boson can also decay to other particles and these decay channels have different branching fractions. 
Higgs boson can decay to a pair of W bosons, Z bosons, photons, bottom quarks, muons and electrons. Higgs boson can also decay to 
Z boson and a photon. It is also interesting to investigate some invisible decay modes of Higgs boson which can be used to put 
upper bounds on dark matter-nucleon scattering cross section like in the Higgs-portal dark matter models.

@@

>>>>

The standard model (SM) of particle physics is so far the best theoretical model to describe the interaction of elementary 
particles using three of the four fundamental forces of nature which are electromagnetic force, strong nuclear force and the weak
nuclear force. Gravitational force is neglected as the strength of this force is very weak at the scales over which elementary 
particle interact with each other. The standard model (SM) of particle physics is divided into two categories, bosonic sector 
and fermionic sector. Bosonic sector contain particles called bosons which mediate the fundamental forces of nature and the
fermionic sector contain particles called fermions which make up all the matter in our universe. SM has three generations of 
matter (fermions) particles.The first generation of fermions consists of up (u) quark, down (d) quark, electron and electron 
neutrino,second generation consist of charm (c) quark, strange (s) quark, muon and muon neutrino and the third generation of matter 
particles has top(t) quark,bottom(b) quark, tau and tau neutrino. The bosonic sector consist of gauge bosons like gluon, photon, 
W^+ W^- Z^0 which mediate strong force, electromagnetic force and weak force respectively. There is one more particle in the 
standard model called the Higgs Boson which gives mass to SM particles via electroweak symmetry breaking mechanism. Higgs boson 
can be produced at the particle colliders like the Large Hadron Collider(LHC) in Geneva,Switzerland.

There are various production modes of Higgs boson like the gluon gluon fusion(ggF), vector boson fusion(VBF), Higgs production 
in association with vector boson (VH, V=W or Z), Higgs production with top quark and top anti-quark(ttH) and Higgs production with 
a single top quark and a quark jet (tHq). Each production channel has its own importance to probe the properties and the coupling 
strength of Higgs boson to fermions and bosons. The main production mode of Higgs bosons at LHC is gluon gluon fusion(ggF). As 
bosons like gluons and photons are massless, they do not interact directly with the Higgs boson and processes like ggF or Higgs 
boson decay to pair of photons are not possible at tree level and can only proceed via loop diagrams which involve W boson or top 
quark in the loop. The Higgs boson can also decay to other particles and these decay channels have different branching fractions. 
Higgs boson can decay to a pair of W bosons, Z bosons, photons, bottom quarks, muons and electrons. Higgs boson can also decay to 
Z boson and a photon. It is also interesting to investigate some invisible decay modes of Higgs boson which can be used to put 
upper bounds on dark matter-nucleon scattering cross section like in the Higgs-portal dark matter models.

@@

The Standard Model (SM) of particle physics has successfully described most of the experimental data till now but a very large number of
free-parameters like fine structure constant \alpha, Weinberg angle or weak mixing angle \theta_W, the coupling constant of strong 
interaction, electroweak symmetry breaking energy scale, Higgs potential self coupling, three mixing angles and the CP-violating phase 
of the CKM matrix which tells us how quarks of different color charge mix with each other, nine yukawa couplings which determine the mass of nine charged fermions and fine tunings related to the origin of masses robustly suggests new physics beyond the SM (BSM).There are lot of specific beyond BSM theories and most of these models involve heavy fields. In order to identify which new physics lies beyond the electroweak (EW) scale, the new parameters of such theories may be constrained by the actual, low energy, experiments. This 
approach requires studying each model individually, and calculating every possible observable. There is another approach immitating 
Fermi’s treatment of beta decay which consists of considering the SM as the first order approximation of the actual theory and by 
completing it by a series of higher dimensional operators. When the electroweak symmetry breaking takes place well below the mass of the 
new particles, the BSM physics is taken into account at the EW scale and below by adding higher dimensional operators to the SM
Lagrangian. They are built out of SM fields and supposed to be invariant under its gauge group. Those operators are the low-energy 
remnant of the high energy theory. This approach does not pretend to guess the complete high-energy model and it is based solely on the 
symmetry of the  theory. The operators are general and the only model dependence is encoded in the size of the operator coefficients, 
which is to be set from experiments. The Higgs boson and fermion (quarks and leptons) coupling will deviate from the standard model 
predictions if these higher dimensional operators are present in low energy effective SM theory.
